Long-term cloud enhancement events of global solar irradiance over Thessaloniki, Greece

### 2024-07-05 article/article.Rmd
Commit 2024-07-05 09:12

For the detection of CE cases we established a baseline of irradiance above which we
characterized each data point as an enhancement event and calculated the over
irradiance (OIR). The OIR is defined as the irradiance difference of the measured
one-minute GHI from the $\text{GHI}_\text{ref}$ corresponding to cloud-free
atmosphere.
<!-- CE identification criterion ($\text{OIR}_i =
\text{GHI}_i - \text{GHI}_\text{CSlim}$), as defined in
Equation\nobreakspace\ref{eq:CE4}. -->
First, we tested two simple approaches for the determination of
$\text{GHI}_\text{ref}$: The Haurwitz's model [@Haurwitz1945], which is a simple
clear sky model and was already adjusted and applied to our data [@Natsis2023], and
the total solar irradiance (TSI) at the top of the atmosphere, adjusted for the
Sun-Earth distance.  We have tested both methods by using an appropriate relative
threshold and/or an additional constant offset.  The initial results showed that we
can detect a big portion of the actual CE events.  However, by inspecting the daily
plots of irradiance it became evident that changes in atmospheric conditions
introduced numerous false positive or false negative results. The main reason for
these discrepancies is the variability of the effects of aerosols and water vapor
which were not taken into account in the two simple methods.  To produce a more
representative reference we included the effects of these factors using a radiative
transfer model (RTM).  The applied methodology is discussed in
section\nobreakspace{}\ref{rtmcs}.

@@

According to @Kazadzis2007, in the period 1997 -- 2005 the mean AOD at $340\,\text{nm}$
is $0.403$ with a trend of $-3.8\pm0.93\,\%$ per year, corresponding to a change of
$0.0153$ per year.  Using an Ångström coefficient $\alpha = 1.6$, this translates to
a change in the Ångström coefficient $\beta$ of $0.00272$ per year (or $\beta=0.084$ in
1997 and $\beta=0.059$ in 2005).  Simulations with uvspec for the above Ångström
coefficients, with WC of
$`r round(BR[year==1997,pw_avg_mm],1)`\,\text{mm}$ and 
$`r round(BR[year==2005,pw_avg_mm],1)`\,\text{mm}$
for 1997 and 2005 respectively, also taken from AERONET, and for a SZA of $55^\circ$
reveal a trend of
$+0.21\,\%$ per year in $\text{GHI}_\text{ref}$. 
The SZA of $55^\circ$ was chosen as representative of all days in the year in order
to get a rough estimate of the annually averaged change in clear sky irradiance.  For
the period 2005 -- 2023 we used the mean monthly values of AOD and WC from AERONET in
a similar simulation scheme to calculate the monthly mean clear-sky irradiance,
and finally the trend of $+0.14\,\%$ per year. We applied these two long-term changes
(see Figure\nobreakspace{}\ref{fig:CS-change}) to the climatological
$\text{GHI}_\text{ref}$, in order to create a more realistic representation of the
clear-sky irradiance for the whole period of study.

```{r CS-change, echo=FALSE, fig.cap="Simulated long-term change in clear sky irradiance relative to the climatological values due to changes in AOD in Thessaloniki for the period 1992 -- 2023."}
knitr::include_graphics("../images/P-CS-change-1.pdf")
```

@@

\begin{figure}% [h!]
        {\centering 
            \subfloat[\label{fig:P-energy-mean}]
                {\includegraphics[width=\linewidth]{../images/P-energy-3} }\\
            \subfloat[\label{fig:P-energy-N}]
                {\includegraphics[width=\linewidth]{../images/P-energy-2} }\\
            \subfloat[\label{fig:P-energy-sum}]
                {\includegraphics[width=\linewidth]{../images/P-energy-1} }
        }
    \caption{Time series for the period 1992 -- 2023 of (a) the yearly CE number of occurrences, (b) the yearly mean OIR and (c) the yearly excess irradiation. The black lines represent the linear trends on the yearly data.}\label{fig:P-energy}
\end{figure}

@@

```{r relative-month-occurrences, echo=FALSE, fig.cap="Seasonal variability of the number of CE events in Thessaloniki for the period 1992 -- 2023 normalized to the maximum occurring in June, in the form of a box and whisker plot. The monthly values have also been normalized to the relative abundance of valid GHI observations. The box contains the data between the lower $25\\,\\%$ and the upper $75\\,\\%$ percentiles, the thick horizontal line and the diamond symbol represent the median and the mean values, respectively. The vertical lines (whiskers) extend between the maximum and minimum monthly values."}
knitr::include_graphics("../images/clim-CE-month-norm-MAX-median-N-MW-1.png")
```

@@

```{r ovir-distribution, echo=FALSE, fig.cap="Relative frequency distribution of CE events in Thessaloniki for the period 1992 -- 2023 as a function of OIR."}
knitr::include_graphics("../images/P-relative-distribution-diff-1.png")
```

@@

```{r ceg-duration-distribution, echo=FALSE, fig.cap="Relative frequency distribution of CE groups of consequent CE events according to their duration for Thessaloniki in the period 1992 -- 2023."}
knitr::include_graphics("../images/groups-1.png")
```

@@

```{r group-2d, echo=FALSE, fig.cap="Relation of excess irradiation of CE groups with their duration for Thessaloniki in the period 1992 -- 2023. The logarithmic color scale denotes the frequency of the respective groups of events."}
knitr::include_graphics("../images/P-groups-bin2d-1.png")
```

@@

```{r P-extreme-distribution, echo=FALSE, fig.cap="Distribution of ECE events in Thessaloniki for the period 1992 -- 2023."}
knitr::include_graphics("../images/P-extreme-distribution-1.png")
```

@@

By creating a clear sky approximation representing the long- and short-term
variability of the expected clear sky GHI, we were able to identify cases of CE
events in Thessaloniki for the period 1992 -- 2023.  After analyzing the CE cases, we
found an increase of
$`r round(yearlyCE[var == "GLB_diff.N", slope], 1)`\pm `r round(2 * yearlyCE[var == "GLB_diff.N", slope.sd], 1)`\,\text{cases}/\text{year}$,
with the mean annual irradiation of the CE events increasing with a rate of
$`r round(yearlyCE[var == "GLB_diff.sum", slope], 1)`\pm `r round(2 * yearlyCE[var == "GLB_diff.sum", slope.sd], 1)`\,\text{kJ}/\text{year}$.
The most active months of CE events are May and June.  We found that continuous CE
events can last up to $140$ minutes, while the duration of $80\,\%$ of them is bellow
$5$ minutes.

>>>>

For the detection of CE cases we established a baseline of irradiance above which we
characterized each data point as an enhancement event and calculated the over
irradiance (OIR). The OIR is defined as the irradiance difference of the measured
one-minute GHI from the $\text{GHI}_\text{ref}$ corresponding to cloud-free
atmosphere.
<!-- CE identification criterion ($\text{OIR}_i =
\text{GHI}_i - \text{GHI}_\text{CSlim}$), as defined in
Equation\nobreakspace\ref{eq:CE4}. -->
First, we tested two simple approaches for the determination of
$\text{GHI}_\text{ref}$: The Haurwitz's model [@Haurwitz1945], which is a simple
clear sky model and was already adjusted and applied to our data [@Natsis2023].
<!--
, and the total solar irradiance (TSI) at the top of the atmosphere, adjusted for the
Sun-Earth distance.
-->
We have tested both methods by using an appropriate relative
threshold and/or an additional constant offset.  The initial results showed that we
can detect a big portion of the actual CE events.  However, by inspecting the daily
plots of irradiance it became evident that changes in atmospheric conditions
introduced numerous false positive or false negative results. The main reason for
these discrepancies is the variability of the effects of aerosols and water vapor
which were not taken into account in the two simple methods.  To produce a more
representative reference we included the effects of these factors using a radiative
transfer model (RTM).  The applied methodology is discussed in
section\nobreakspace{}\ref{rtmcs}.

@@

According to @Kazadzis2007, in the period 1997 -- 2005 the mean AOD at $340\,\text{nm}$
is $0.403$ with a trend of $-3.8\pm0.93\,\%$ per year, corresponding to a change of
$0.0153$ per year.  Using an Ångström coefficient $\alpha = 1.6$, this translates to
a change in the Ångström coefficient $\beta$ of $0.00272$ per year (or $\beta=0.084$ in
1997 and $\beta=0.059$ in 2005).  Simulations with uvspec for the above Ångström
coefficients, with WC of
$`r round(BR[year==1997,pw_avg_mm],1)`\,\text{mm}$ and
$`r round(BR[year==2005,pw_avg_mm],1)`\,\text{mm}$
for 1997 and 2005 respectively, where for 1997 we used the first available WC data
(i.e. 2003) as a proxy, also taken from AERONET, and for a SZA of $55^\circ$ reveal a
trend of
$+0.21\,\%$ per year in $\text{GHI}_\text{ref}$.
The SZA of $55^\circ$ was chosen as representative of all days in the year in order
to get a rough estimate of the annually averaged change in clear sky irradiance.  For
the period 2005 -- 2023 we used the mean monthly values of AOD and WC from AERONET in
a similar simulation scheme to calculate the monthly mean clear-sky irradiance,
and finally the trend of $+0.14\,\%$ per year. We applied these two long-term changes
(see Figure\nobreakspace{}\ref{fig:CS-change}) to the climatological
$\text{GHI}_\text{ref}$, in order to create a more realistic representation of the
clear-sky irradiance for the whole period of study.

```{r CS-change, echo=FALSE, fig.cap="Simulated long-term change in clear sky irradiance relative to the climatological values due to changes in AOD in Thessaloniki for the period 1993 -- 2023."}
knitr::include_graphics("../images/P-CS-change-1.pdf")
```

@@

\begin{figure}% [h!]
        {\centering 
            \subfloat[\label{fig:P-energy-mean}]
                {\includegraphics[width=\linewidth]{../images/P-energy-3} }\\
            \subfloat[\label{fig:P-energy-N}]
                {\includegraphics[width=\linewidth]{../images/P-energy-2} }\\
            \subfloat[\label{fig:P-energy-sum}]
                {\includegraphics[width=\linewidth]{../images/P-energy-1} }
        }
    \caption{Time series for the period 1993 -- 2023 of (a) the yearly CE number of occurrences, (b) the yearly mean OIR and (c) the yearly excess irradiation. The black lines represent the linear trends on the yearly data.}\label{fig:P-energy}
\end{figure}

@@

```{r relative-month-occurrences, echo=FALSE, fig.cap="Seasonal variability of the number of CE events in Thessaloniki for the period 1993 -- 2023 normalized to the maximum occurring in June, in the form of a box and whisker plot. The monthly values have also been normalized to the relative abundance of valid GHI observations. The box contains the data between the lower $25\\,\\%$ and the upper $75\\,\\%$ percentiles, the thick horizontal line and the diamond symbol represent the median and the mean values, respectively. The vertical lines (whiskers) extend between the maximum and minimum monthly values."}
knitr::include_graphics("../images/clim-CE-month-norm-MAX-median-N-MW-1.png")
```

@@

```{r ovir-distribution, echo=FALSE, fig.cap="Relative frequency distribution of CE events in Thessaloniki for the period 1993 -- 2023 as a function of OIR."}
knitr::include_graphics("../images/P-relative-distribution-diff-1.png")
```

@@

```{r ceg-duration-distribution, echo=FALSE, fig.cap="Relative frequency distribution of CE groups of consequent CE events according to their duration for Thessaloniki in the period 1993 -- 2023."}
knitr::include_graphics("../images/groups-1.png")
```

@@

```{r group-2d, echo=FALSE, fig.cap="Relation of excess irradiation of CE groups with their duration for Thessaloniki in the period 1993 -- 2023. The logarithmic color scale denotes the frequency of the respective groups of events."}
knitr::include_graphics("../images/P-groups-bin2d-1.png")
```

@@

```{r P-extreme-distribution, echo=FALSE, fig.cap="Distribution of ECE events in Thessaloniki for the period 1993 -- 2023."}
knitr::include_graphics("../images/P-extreme-distribution-1.png")
```

@@

By creating a clear sky approximation representing the long- and short-term
variability of the expected clear sky GHI, we were able to identify cases of CE
events in Thessaloniki for the period 1993 -- 2023.  After analyzing the CE cases, we
found an increase of
$`r round(yearlyCE[var == "GLB_diff.N", slope], 1)`\pm `r round(2 * yearlyCE[var == "GLB_diff.N", slope.sd], 1)`\,\text{cases}/\text{year}$,
with the mean annual irradiation of the CE events increasing with a rate of
$`r round(yearlyCE[var == "GLB_diff.sum", slope], 1)`\pm `r round(2 * yearlyCE[var == "GLB_diff.sum", slope.sd], 1)`\,\text{kJ}/\text{year}$.
The most active months of CE events are May and June.  We found that continuous CE
events can last up to $140$ minutes, while the duration of $80\,\%$ of them is bellow
$5$ minutes.

### 2024-07-05 article/article.tex
Commit 2024-07-05 09:12

For the detection of CE cases we established a baseline of irradiance
above which we characterized each data point as an enhancement event and
calculated the over irradiance (OIR). The OIR is defined as the
irradiance difference of the measured one-minute GHI from the
\(\text{GHI}_\text{ref}\) corresponding to cloud-free atmosphere. First,
we tested two simple approaches for the determination of
\(\text{GHI}_\text{ref}\): The Haurwitz's model \citep{Haurwitz1945},
which is a simple clear sky model and was already adjusted and applied
to our data \citep{Natsis2023}, and the total solar irradiance (TSI) at
the top of the atmosphere, adjusted for the Sun-Earth distance. We have
tested both methods by using an appropriate relative threshold and/or an
additional constant offset. The initial results showed that we can
detect a big portion of the actual CE events. However, by inspecting the
daily plots of irradiance it became evident that changes in atmospheric
conditions introduced numerous false positive or false negative results.
The main reason for these discrepancies is the variability of the
effects of aerosols and water vapor which were not taken into account in
the two simple methods. To produce a more representative reference we
included the effects of these factors using a radiative transfer model
(RTM). The applied methodology is discussed in
section\nobreakspace{}\ref{rtmcs}.

@@

According to \citet{Kazadzis2007}, in the period 1997 -- 2005 the mean
AOD at \(340\,\text{nm}\) is \(0.403\) with a trend of
\(-3.8\pm0.93\,\%\) per year, corresponding to a change of \(0.0153\)
per year. Using an Ångström coefficient \(\alpha = 1.6\), this
translates to a change in the Ångström coefficient \(\beta\) of
\(0.00272\) per year (or \(\beta=0.084\) in 1997 and \(\beta=0.059\) in
2005). Simulations with uvspec for the above Ångström coefficients, with
WC of \(15.7\,\text{mm}\) and \(15.4\,\text{mm}\) for 1997 and 2005
respectively, also taken from AERONET, and for a SZA of \(55^\circ\)
reveal a trend of \(+0.21\,\%\) per year in \(\text{GHI}_\text{ref}\).
The SZA of \(55^\circ\) was chosen as representative of all days in the
year in order to get a rough estimate of the annually averaged change in
clear sky irradiance. For the period 2005 -- 2023 we used the mean
monthly values of AOD and WC from AERONET in a similar simulation scheme
to calculate the monthly mean clear-sky irradiance, and finally the
trend of \(+0.14\,\%\) per year. We applied these two long-term changes
(see Figure\nobreakspace{}\ref{fig:CS-change}) to the climatological
\(\text{GHI}_\text{ref}\), in order to create a more realistic
representation of the clear-sky irradiance for the whole period of
study.

@@

\caption{Simulated long-term change in clear sky irradiance relative to the climatological values due to changes in AOD in Thessaloniki for the period 1992 -- 2023.}\label{fig:CS-change}
\end{figure}

@@

\caption{Seasonal variability of the number of CE events in Thessaloniki for the period 1992 -- 2023 normalized to the maximum occurring in June, in the form of a box and whisker plot. The monthly values have also been normalized to the relative abundance of valid GHI observations. The box contains the data between the lower $25\,\%$ and the upper $75\,\%$ percentiles, the thick horizontal line and the diamond symbol represent the median and the mean values, respectively. The vertical lines (whiskers) extend between the maximum and minimum monthly values.}\label{fig:relative-month-occurrences}
\end{figure}

@@

\caption{Relative frequency distribution of CE events in Thessaloniki for the period 1992 -- 2023 as a function of OIR.}\label{fig:ovir-distribution}
\end{figure}

@@

\caption{Relative frequency distribution of CE groups of consequent CE events according to their duration for Thessaloniki in the period 1992 -- 2023.}\label{fig:ceg-duration-distribution}
\end{figure}

@@

\caption{Relation of excess irradiation of CE groups with their duration for Thessaloniki in the period 1992 -- 2023. The logarithmic color scale denotes the frequency of the respective groups of events.}\label{fig:group-2d}
\end{figure}

@@

\caption{Distribution of ECE events in Thessaloniki for the period 1992 -- 2023.}\label{fig:P-extreme-distribution}
\end{figure}

@@

By creating a clear sky approximation representing the long- and
short-term variability of the expected clear sky GHI, we were able to
identify cases of CE events in Thessaloniki for the period 1992 -- 2023.
After analyzing the CE cases, we found an increase of
\(45.6\pm 21.9\,\text{cases}/\text{year}\), with the mean annual
irradiation of the CE events increasing with a rate of
\(116.9\pm 67.8\,\text{kJ}/\text{year}\). The most active months of CE
events are May and June. We found that continuous CE events can last up
to \(140\) minutes, while the duration of \(80\,\%\) of them is bellow
\(5\) minutes.

>>>>

For the detection of CE cases we established a baseline of irradiance
above which we characterized each data point as an enhancement event and
calculated the over irradiance (OIR). The OIR is defined as the
irradiance difference of the measured one-minute GHI from the
\(\text{GHI}_\text{ref}\) corresponding to cloud-free atmosphere. First,
we tested two simple approaches for the determination of
\(\text{GHI}_\text{ref}\): The Haurwitz's model \citep{Haurwitz1945},
which is a simple clear sky model and was already adjusted and applied
to our data \citep{Natsis2023}. We have tested both methods by using an
appropriate relative threshold and/or an additional constant offset. The
initial results showed that we can detect a big portion of the actual CE
events. However, by inspecting the daily plots of irradiance it became
evident that changes in atmospheric conditions introduced numerous false
positive or false negative results. The main reason for these
discrepancies is the variability of the effects of aerosols and water
vapor which were not taken into account in the two simple methods. To
produce a more representative reference we included the effects of these
factors using a radiative transfer model (RTM). The applied methodology
is discussed in section\nobreakspace{}\ref{rtmcs}.

@@

According to \citet{Kazadzis2007}, in the period 1997 -- 2005 the mean
AOD at \(340\,\text{nm}\) is \(0.403\) with a trend of
\(-3.8\pm0.93\,\%\) per year, corresponding to a change of \(0.0153\)
per year. Using an Ångström coefficient \(\alpha = 1.6\), this
translates to a change in the Ångström coefficient \(\beta\) of
\(0.00272\) per year (or \(\beta=0.084\) in 1997 and \(\beta=0.059\) in
2005). Simulations with uvspec for the above Ångström coefficients, with
WC of \(15.7\,\text{mm}\) and \(15.4\,\text{mm}\) for 1997 and 2005
respectively, where for 1997 we used the first available WC data
(i.e.~2003) as a proxy, also taken from AERONET, and for a SZA of
\(55^\circ\) reveal a trend of \(+0.21\,\%\) per year in
\(\text{GHI}_\text{ref}\). The SZA of \(55^\circ\) was chosen as
representative of all days in the year in order to get a rough estimate
of the annually averaged change in clear sky irradiance. For the period
2005 -- 2023 we used the mean monthly values of AOD and WC from AERONET
in a similar simulation scheme to calculate the monthly mean clear-sky
irradiance, and finally the trend of \(+0.14\,\%\) per year. We applied
these two long-term changes (see
Figure\nobreakspace{}\ref{fig:CS-change}) to the climatological
\(\text{GHI}_\text{ref}\), in order to create a more realistic
representation of the clear-sky irradiance for the whole period of
study.

@@

\caption{Simulated long-term change in clear sky irradiance relative to the climatological values due to changes in AOD in Thessaloniki for the period 1993 -- 2023.}\label{fig:CS-change}
\end{figure}

@@

\caption{Seasonal variability of the number of CE events in Thessaloniki for the period 1993 -- 2023 normalized to the maximum occurring in June, in the form of a box and whisker plot. The monthly values have also been normalized to the relative abundance of valid GHI observations. The box contains the data between the lower $25\,\%$ and the upper $75\,\%$ percentiles, the thick horizontal line and the diamond symbol represent the median and the mean values, respectively. The vertical lines (whiskers) extend between the maximum and minimum monthly values.}\label{fig:relative-month-occurrences}
\end{figure}

@@

\caption{Relative frequency distribution of CE events in Thessaloniki for the period 1993 -- 2023 as a function of OIR.}\label{fig:ovir-distribution}
\end{figure}

@@

\caption{Relative frequency distribution of CE groups of consequent CE events according to their duration for Thessaloniki in the period 1993 -- 2023.}\label{fig:ceg-duration-distribution}
\end{figure}

@@

\caption{Relation of excess irradiation of CE groups with their duration for Thessaloniki in the period 1993 -- 2023. The logarithmic color scale denotes the frequency of the respective groups of events.}\label{fig:group-2d}
\end{figure}

@@

\caption{Distribution of ECE events in Thessaloniki for the period 1993 -- 2023.}\label{fig:P-extreme-distribution}
\end{figure}

@@

By creating a clear sky approximation representing the long- and
short-term variability of the expected clear sky GHI, we were able to
identify cases of CE events in Thessaloniki for the period 1993 -- 2023.
After analyzing the CE cases, we found an increase of
\(45.6\pm 21.9\,\text{cases}/\text{year}\), with the mean annual
irradiation of the CE events increasing with a rate of
\(116.9\pm 67.8\,\text{kJ}/\text{year}\). The most active months of CE
events are May and June. We found that continuous CE events can last up
to \(140\) minutes, while the duration of \(80\,\%\) of them is bellow
\(5\) minutes.
